\subsection{Monotone functions}

32
33

\begin{exercise}{32}
Deduce from Theorem 2.17 that a monotone function $f:\R\to\R$ has points of continuity in every open interval.
\end{exercise}
\begin{proof}
We know that every open interval in the real line is not countable (because we can find a bijection between $(a,b)$ and $\R$. Hence, if there existed an open interval without a point of continuity, then the monotone function would have uncountably many points of discontinuity, which contradicts Theorem 2.17.
\end{proof} 

\begin{exercise}{33}
Let $f:[a,b]\to\R$ be monotone. Given $n$ distinct points $a<x_1<x_2<\dots<x_n<b$, show that $\sum_{i=1}^n\absoluteValue{f(x_i+)-f(x_i-)}\leq \absoluteValue{f(b)-f(a)}$. Use this to give another proof that $f$ has at most countably many (jump) discontinuities.
\end{exercise}
\begin{proof}
Without loss of generality, suppose $f$ is monotonically increasing. We will prove $\sum_{i=1}^n\absoluteValue{f(x_i+)-f(x_i-)}\leq \absoluteValue{f(b)-f(a)}$ by induction.

Base case: from monotonicity, we have that for any point $x_i$, $f(x_i+)-f(x_i-)\leq f(b)-f(a)$ (as otherwise $f$ wouldn't be monotonous or it wouldn't be the case that $a<x_i<b$).

Suppose the statement holds for $i=k$. Then $\sum_{i=1}^k\absoluteValue{f(x_i+)-f(x_i-)}\leq \absoluteValue{f(b)-f(a)}$. We will use the same argument as above. By monotonicity we have that $f(x_{k+1}+)-f(x_{k+1}-)\leq f(b)-f(x_k+)$. Now, since $\sum_{i=1}^k f(x_i+)-f(x_i-)\leq f(x_{k+1}+)-f(a)$ (by the induction hypothesis, and the fact that $x_k<x_{k+1}$). Adding those two together gives us the desired result.

If there were are an uncountable number of discontinuities, then we could not enumerate the elements of the domain where there the discountinuities and the inequality would not hold.
\end{proof} 